preciseTAD: A machine learning framework for precise 3D domain boundary prediction at base-level resolution

### 2020-09-30 BMC_Genome_Biology_submission/cover_letter_Genome_Biology.Rmd
final touches

Our manuscript is submitted for the "Software" section. The main highlights of the software are:

@@

- *preciseTAD* prioritizes genomic signatures associated with domain boundaries. Trained on any 3D genomic features (e.g., besides TAD/loop boundaries, regions of A/B compartment switching, frequently interacting regions (FIREs), differentially interacting chromatin regions, etc.), *preciseTAD* defines (epi)genomic signatures most predictive of those features.

@@

- Boundaries of TADs and loops are associated with the same genome annotation signatures, suggesting similar mechanisms of their formation.

- Biological properties of domain boundaries predicted using *preciseTAD* include the highest genomic signal of CTCF and other transcription factors, closer proximity to other genomic signatures, and more conservation across cell lines.

@@

In summary, our systematic machine learning framework evaluates previously overlooked properties of 3D interactions and 1D regulatory data, linking them into the most optimal model for domain boundary predictions. This improved precision in the domain boundary location can provide deeper insight into the association between genomic regulators and the 3D genome organization.

>>>>

Our manuscript is submitted under the "Software" category. The main highlights of our software are:

@@

- *preciseTAD* prioritizes regulatory elements associated with domain boundaries. Trained on any 3D features (e.g., besides TAD/loop boundaries, regions of A/B compartment switching, frequently interacting regions (FIREs), differentially interacting chromatin regions, etc.), *preciseTAD* defines (epi)genomic signatures most predictive of those features.

@@

- Boundaries of TADs and loops are associated with the same regulatory signatures, suggesting similar mechanisms of their formation.

- Biological properties of domain boundaries predicted using *preciseTAD* include the highest genomic signal of CTCF and other transcription factors, closer proximity to other genomic signatures, and higher` conservation across cell lines.

@@

In summary, our systematic machine learning framework evaluates previously overlooked properties of 3D interactions and 1D regulatory data, linking them into the most optimal model for domain boundary predictions. This improved precision in the domain boundary location will provide deeper insight into the association between genomic regulators and the 3D genome organization.

### 2020-09-30 Manuscript/manuscript.Rmd
final touches

Guided by both low-resolution Hi-C data and high-resolution genome annotation data, *preciseTAD* predicts base-level resolution boundaries, alleviating resolution limitations of Hi-C data. However, a natural question is how resolution (width) of boundary regions identified by *preciseTAD* (PTBRs) compares with that of Hi-C data. Our preliminary observations indicate that, under most optimal settings, the width of PTBRs parallels the resolution of Hi-C data. Furthermore, each PTBR is formed by several sub-regions with the probability of being a boundary defined by the threshold $t$ ($t=1$ in the current study). Yet, the preciseTAD boundary points (PTBPs, medoids identified within each PTBR) had the highest density of CTCF and other transcription factor binding signal (Figure 5). Our results are in line with the emergent view that domain boundaries are dynamic [@chang2020tads], and their well-defined location arises as a consequence of the population average in bulk Hi-C data [@fudenberg2016formation; @Flyamer:2017aa].

>>>>

Guided by both low-resolution Hi-C data and high-resolution genome annotation data, *preciseTAD* predicts base-level resolution boundaries, alleviating resolution limitations of Hi-C data. However, a natural question is how resolution (width) of boundary regions identified by *preciseTAD* (PTBRs) compares with that of Hi-C data. Our preliminary observations indicate that, under most optimal settings, the width of PTBRs parallels the resolution of Hi-C data. Furthermore, each PTBR is formed by several sub-regions with the probability of being a boundary defined by the threshold $t$ ($t=1$ in the current study). Yet, the preciseTAD boundary points (PTBPs, medoids identified within each PTBR) had the highest density of CTCF and other transcription factor binding signal (Figure 5). Our results are in line with the emergent view that domain boundaries are flexible [@chang2020tads; @Maeshima:2020aa], and their well-defined location arises as a consequence of the population average in bulk Hi-C data [@fudenberg2016formation; @Flyamer:2017aa].
